Longitudinal Principal Manifold Estimation

### 2023-07-05 paper/main.tex
update simulation description

To assess the performance of the LPME algorithm, we compare its performance with the that of two alternative approaches on simulated datasets. We consider settings in which $d = 1$ and $D = 2$, $d = 1$ and $D = 3$, and $d = 2$ and $D = 3$, with several manifolds being used to generate datasets in each setting. For each manifold, differing values for the time duration, interval between observations, noise levels within and between time points, and the type and magnitude of structural changes in the underlying manifold over time were considered. The embedding functions used to define each manifold are given in Table \ref{table:1}. In these functions, the value of $\zeta$ and the functional form of $g(\cdot)$ are varied between trials. The values of $\iota$ are randomly sampled from a normal distribution with mean zero and variance that is varied between trials. Meanwhile, in cases where $d=1$, $\alpha$ and $\beta$ are scalars that are set externally. If $d>1$, then $\alpha$ and $\beta$ take the form of vectors of length $d$, with the values drawn from a normal distribution with mean equal to 1 and variance that is varied between trials.

\begin{table}
  \centering
  \begin{tabular}{|c c c c c|}
    \hline
    Case & $d$ & $D$ & $f(\mathbf{r})$ & Domain \\
    \hline
    1 & 1 & 2 & $\left(r, \ \alpha \text{sin} \ (\beta r + \frac{\pi}{2})\right) + \zeta\mathbf{g}(t) + \iota$ & $-3 \leq r \leq 3$ \\
    2 & 1 & 2 & $\left(r, \ \alpha\text{sin} \ (\beta r)\right) + \zeta\mathbf{g}(t) + \iota$ & $-3\pi \leq r \leq 3\pi$ \\
    3 & 1 & 2 & $\left(\alpha_1 \text{cos} \ (\beta_1 r), \ \alpha_2\text{cos} \ (\beta_2 r)\right) + \zeta\mathbf{g}(t) + \iota$ & $-\frac{4\pi}{5} \leq r \leq \frac{\pi}{2}$ \\
    4 & 1 & 3 & $\left(r, (\alpha_1r + \beta_1)^2, \ (\alpha_2r + \beta_2)^3\right) + \zeta\mathbf{g}(t) + \iota$ & $-1 \leq r \leq 1$ \\
    5 & 1 & 3 & $\left(r, \ \alpha_1\text{cos} \ (\beta_1 r), \ \alpha_2\text{sin} \ (\beta_2 r) \right) + \zeta\mathbf{g}(t) + \iota$ & $0 \leq r \leq 3\pi$ \\
    6 & 2 & 3 & $\left(\beta_1r_1, \ \beta_2r_2, \ \alpha_1(\alpha_2\|\beta\mathbf{r}\|^2)\right) + \zeta\mathbf{g}(t) + \iota$ & $-1 \leq r_1, r_2 \leq 1$\\
    7 & 2 & 3 & $\left(\alpha_1\beta_1r_1\text{cos} \ (\alpha_1r_1), \ \alpha_2\beta_2r_1\text{sin} \ (\alpha_2r_1), \ r_2\right) + \zeta\mathbf{g}(t) + \iota$ & $0 \leq r_1 \leq 3\pi$; \ $-1 \leq r_2 \leq 1$\\
    8 & 2 & 3 & $\left(\alpha_1\text{sin}(\beta_1r_1)\text{cos}(\beta_2r_2), \ \alpha_1\text{sin}(\beta_1r_1)\text{sin}(\beta_2r_2), \ \alpha_1\text{cos}(\beta_1r_1)\right) + \zeta\mathbf{g}(t) + \iota$ & $0 \leq r_1 \leq \pi$; \ $0 \leq r_2 \leq 2\pi$\\
    9 & 2 & 3 & $\left(\theta\text{sin} \ (\beta_1r_1)\text{cos} \ (\beta_2r_2), \ \theta\text{sin} \ (\beta_1)\text{sin} \ (\beta_2r_2), \ \theta\text{cos} \ (\beta_1r_1)\right) + \zeta\mathbf{g}(t) + \iota$ & $0 \leq r_1 \leq \pi$; \ $0 \leq r_2 \leq 2\pi$\\
    \hline
  \end{tabular}
  \caption{Embedding functions used for simulation trials.}
  \label{table:1}
\end{table}

>>>>

To assess the performance of the LPME algorithm, we compare its performance with the that of two alternative approaches on simulated datasets. We consider settings in which $d = 1$ and $D = 2$, $d = 1$ and $D = 3$, and $d = 2$ and $D = 3$, with several manifolds being used to generate datasets in each setting. For each manifold, differing values for the time duration, interval between observations, noise levels within and between time points, and the type and magnitude of structural changes in the underlying manifold over time were considered.

The embedding functions used to define each manifold are given in Table \ref{table:1}. In these functions, the value of $\zeta$ and the functional form of $g(\cdot)$ are varied between trials and represent change in the underlying manifold over time. The function $g(\cdot)$ may be constant, linear, quadratic, or sinusoidal, with $\zeta$ serving as a scalar multiplier. The values of $\iota$ represent noise in the high-dimensional space, and are randomly sampled from a normal distribution with mean zero and variance that is varied between trials. Meanwhile, $\alpha$ and $\beta$ are vectors of length $d$ that are drawn from a normal distribution with mean equal to 1 and variance that is varied between trials. These values describe random fluctuations in the manifold between time points.

\begin{table}
  \centering
  \begin{tabular}{|c c c c c|}
    \hline
    Case & $d$ & $D$ & $f_t(\mathbf{r})$ & Domain \\
    \hline
    1 & 1 & 2 & $\left(r, \ \alpha \text{sin} \ (\beta r + \frac{\pi}{2})\right) + \zeta\mathbf{g}(t) + \iota$ & $-3 \leq r \leq 3$ \\
    2 & 1 & 2 & $\left(r, \ \alpha\text{sin} \ (\beta r)\right) + \zeta\mathbf{g}(t) + \iota$ & $-3\pi \leq r \leq 3\pi$ \\
    3 & 1 & 2 & $\left(\alpha_1 \text{cos} \ (\beta_1 r), \ \alpha_2\text{cos} \ (\beta_2 r)\right) + \zeta\mathbf{g}(t) + \iota$ & $-\frac{4\pi}{5} \leq r \leq \frac{\pi}{2}$ \\
    4 & 1 & 3 & $\left(r, (\alpha_1r + \beta_1)^2, \ (\alpha_2r + \beta_2)^3\right) + \zeta\mathbf{g}(t) + \iota$ & $-1 \leq r \leq 1$ \\
    5 & 1 & 3 & $\left(r, \ \alpha_1\text{cos} \ (\beta_1 r), \ \alpha_2\text{sin} \ (\beta_2 r) \right) + \zeta\mathbf{g}(t) + \iota$ & $0 \leq r \leq 3\pi$ \\
    6 & 2 & 3 & $\left(\beta_1r_1, \ \beta_2r_2, \ \alpha_1(\alpha_2\|\beta\mathbf{r}\|^2)\right) + \zeta\mathbf{g}(t) + \iota$ & $-1 \leq r_1, r_2 \leq 1$\\
    7 & 2 & 3 & $\left(\alpha_1\beta_1r_1\text{cos} \ (\alpha_1r_1), \ \alpha_2\beta_2r_1\text{sin} \ (\alpha_2r_1), \ r_2\right) + \zeta\mathbf{g}(t) + \iota$ & $0 \leq r_1 \leq 3\pi$; \ $-1 \leq r_2 \leq 1$\\
    8 & 2 & 3 & $\left(\alpha_1\text{sin}(\beta_1r_1)\text{cos}(\beta_2r_2), \ \alpha_1\text{sin}(\beta_1r_1)\text{sin}(\beta_2r_2), \ \alpha_1\text{cos}(\beta_1r_1)\right) + \zeta\mathbf{g}(t) + \iota$ & $0 \leq r_1 \leq \pi$; \ $0 \leq r_2 \leq 2\pi$\\
    9 & 2 & 3 & $\left(\theta\text{sin} \ (\beta_1r_1)\text{cos} \ (\beta_2r_2), \ \theta\text{sin} \ (\beta_1)\text{sin} \ (\beta_2r_2), \ \theta\text{cos} \ (\beta_1r_1)\right) + \zeta\mathbf{g}(t) + \iota$ & $0 \leq r_1 \leq \pi$; \ $0 \leq r_2 \leq 2\pi$\\
    \hline
  \end{tabular}
  \caption{Embedding functions used for simulation trials.}
  \label{table:1}
\end{table}
